Romans 1 by 1. Documenting a population database for the Roman world

### 2016-02-10 source/10/Paper proposal for EAGLE final.tex
orfans and widows
percorsi tutti i file per eliminare orfane e vedove inserendo \newpage
o \vspace per spostare leggermente i paragrafi

A second issue in the study of palaeography involves the access and publication of epigraphic materials. Access to
large corpora of inscriptions has traditionally been limited to a small audience of scholars and site visitors. Apart
from a few museum collections, which happen to have inscriptions arranged in roughly chronological order,\footnote{
Perhaps the finest example of a museum in which one can gain and understanding of palaeography (and inscriptions as
whole) is the Museo Epigrafico Nazionale, Roma. } there are few places where one can visually experience the
development of carving styles over time. For the lucky few who attain access (and permission) to study a large corpus
of inscriptions, publishing these texts with supporting images can be challenging and expensive. Studies of
palaeography in ancient inscriptions have often been, by necessity, selective with images making it difficult for both
the author and readers to develop a detailed understanding of carving trends and practices. In this traditional format,
inscriptions were also separated from their archaeological context and the accompanying artwork while the reader,
viewing only the lettering, was often removed from the visual elements of the inscription (e.g. the type of stone, use
of spacing and decoration, letter size). This is a suboptimal way of assessing epigraphic evidence.

The advent of online corpora have increased both the access and the development of discussions regarding epigraphic
monumentality, including new methodologies, approaches as well as attempts to redefine genre classifications and
terminology.\footnote{\citet[22-39]{woolf1996}, \citet{eck2010}, \citet[1-10]{panciera2012}, \citet[1-17]{graham2013}. For a recent collection and analysis of epigraphic databases, see \citet[78-85]{Elliot2015}.} A
number of recent studies have used changes in the appearance of inscriptions over time, such as the use of different
media, decorative and paratextual elements\footnote{On marks in the text see \citet[26]{susini1973} and \citet[143-155]{cooley2014}.} (e.g. ivy leaves as interpuncts (hederae distinguentes),\footnote{\citet[293-303]{hommel1970}}
abbreviations\footnote{\citet[111]{Gordon1957}} and spacing between words\footnote{On using
space to reconstruct inscriptions see \citet[195-226]{alfoldy1995}, \citet[151-176]{grasby2002}.} alongside lettering, as dating
criteria.\footnote{ For an extensive list of dating conventions cf. \citet[Chapter 20]{Manzella1987}. }
Panciera’s recent article, in particular, advocates the significance of the public context and visibility of
inscriptions.\footnote{ ``I would propose to regard as an ‘inscription’ any particular type of written human
communication of the sort that we would today call unidirectional…not being addressed to a person or to a group but to
a collectivity, and for this reason is made with the location, writing technique, graphic form and impagination, mode
and register of expression chosen because they are most suitable to the attainment of its intended goal.'' \citep[8]{panciera2012}.} In this vein, it is worth considering how the physical characteristics of an inscription belong within a
broader assessment of a culture of writing. Is palaeography the study of letterforms alone? Was a focus on letterforms
a deliberate choice or a product of the traditional constraints in accessibility and publication of epigraphic
materials? The following case study from Aphrodisias will explore these questions further.

@@

\section{Publishing inscriptions: A Case Study of Column Dedications from The Temple Of Aphrodite at Aphrodisias}

@@

The two published texts reveal similar inscriptions with a few variations.\footnote{ MAMA 437 and 438 are represented in
their digital format as published the Packhum website \ (Aphrodisias 108 and 109 (respectively)).
\url{http://epigraphy.packhum.org/inscriptions/main?url=oi\%3Fikey\%3D256987}} IAph 1.4 (MAMA 437) has a slightly less
impressive ordinatio, particularly in line 4 with an odd line break, a misspelling of Ammias (as Amias) and Adrastus
(line 5) with a sigma omitted. It is difficult to determine how or why these errors occurred without indications of
spacing or an examination of the stone. \ One could argue that spelling and arrangement of the text were not important
to a broader audience at Aphrodisias, however, this is negated by a second copy of the dedication (IAph 1.5, MAMA 438),
which portrays a more skillful execution and arrangement of the text. Spelling errors are rectified and the arrangement
of the text creates a distinction in Line 5 between Eumachus and Ammias, and for the demos, which is isolated on the
last line. These subtle variations only emerge with a close reading of the text, and it would be easy to overlook them.
The published format of the texts also makes them appear similar in lettering and arrangement (e.g. left indentation).

@@

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{PaperproposalforEAGLEfinal-img001.png}
\caption{Screenshot of IAph 4 webpage}
\label{fig:1}
\end{figure}

@@

While traditional text-based modes of publication such as MAMA VIII have, at times, constrained the study of hands to an
analysis of lettering, digital corpora offer new opportunities to view, analyse and incorporate broader scope of visual
elements in the interpretation of an inscription. This is not to say that the traditional methods or charts of
letterforms development should be abandoned, merely that the methodology could be expanded, as it has been in this case
study, to include aspects of textual organization (spacing, use of decoration and punctuation, letter size). The
question arises: how do we achieve this through the website materials?

>>>>

A second issue in the study of palaeography involves the access and publication of epigraphic materials. Access to
large corpora of inscriptions has traditionally been limited to a small audience of scholars and site visitors. Apart
from a few museum collections, which happen to have inscriptions arranged in roughly chronological order,\footnote{
Perhaps the finest example of a museum in which one can gain and understanding of palaeography (and inscriptions as
whole) is the Museo Epigrafico Nazionale, Roma. } there are few places where one can visually experience the
development of carving styles over time. For the lucky few who attain access (and permission) to study a large corpus
of inscriptions, publishing these texts with supporting images can be challenging and expensive. Studies of
palaeography in ancient inscriptions have often been, by necessity, selective with images making it difficult for both
the author and readers to develop a detailed understanding of carving trends and practices. In this traditional format,
inscriptions were also separated from their archaeological context and the accompanying artwork while the reader,
viewing only the lettering, was often removed from the visual elements of the inscription (e.g. the type of stone, use
of spacing and decoration, letter size). This is a suboptimal way of assessing epigraphic evidence. 
\vspace{5pt}
The advent of online corpora have increased both the access and the development of discussions regarding epigraphic
monumentality, including new methodologies, approaches as well as attempts to redefine genre classifications and
terminology.\footnote{\citet[22-39]{woolf1996}, \citet{eck2010}, \citet[1-10]{panciera2012}, \citet[1-17]{graham2013}. For a recent collection and analysis of epigraphic databases, see \citet[78-85]{Elliot2015}.} A
number of recent studies have used changes in the appearance of inscriptions over time, such as the use of different
media, decorative and paratextual elements\footnote{On marks in the text see \citet[26]{susini1973} and \citet[143-155]{cooley2014}.} (e.g. ivy leaves as interpuncts (hederae distinguentes),\footnote{\citet[293-303]{hommel1970}}
abbreviations\footnote{\citet[111]{Gordon1957}} and spacing between words\footnote{On using
space to reconstruct inscriptions see \citet[195-226]{alfoldy1995}, \citet[151-176]{grasby2002}.} alongside lettering, as dating
criteria.\footnote{ For an extensive list of dating conventions cf. \citet[Chapter 20]{Manzella1987}. }
Panciera’s recent article, in particular, advocates the significance of the public context and visibility of
inscriptions.\footnote{ ``I would propose to regard as an ‘inscription’ any particular type of written human
communication of the sort that we would today call unidirectional…not being addressed to a person or to a group but to
a collectivity, and for this reason is made with the location, writing technique, graphic form and impagination, mode
and register of expression chosen because they are most suitable to the attainment of its intended goal.'' \citep[8]{panciera2012}.} In this vein, it is worth considering how the physical characteristics of an inscription belong within a
broader assessment of a culture of writing. Is palaeography the study of letterforms alone? Was a focus on letterforms
a deliberate choice or a product of the traditional constraints in accessibility and publication of epigraphic
materials? The following case study from Aphrodisias will explore these questions further.

@@

\vspace{5pt}
\section{Publishing inscriptions: A Case Study of Column Dedications from The Temple Of Aphrodite at Aphrodisias}

@@

\newpage
The two published texts reveal similar inscriptions with a few variations.\footnote{ MAMA 437 and 438 are represented in
their digital format as published the Packhum website \ (Aphrodisias 108 and 109 (respectively)).
\url{http://epigraphy.packhum.org/inscriptions/main?url=oi\%3Fikey\%3D256987}} IAph 1.4 (MAMA 437) has a slightly less
impressive ordinatio, particularly in line 4 with an odd line break, a misspelling of Ammias (as Amias) and Adrastus
(line 5) with a sigma omitted. It is difficult to determine how or why these errors occurred without indications of
spacing or an examination of the stone. \ One could argue that spelling and arrangement of the text were not important
to a broader audience at Aphrodisias, however, this is negated by a second copy of the dedication (IAph 1.5, MAMA 438),
which portrays a more skillful execution and arrangement of the text. Spelling errors are rectified and the arrangement
of the text creates a distinction in Line 5 between Eumachus and Ammias, and for the demos, which is isolated on the
last line. These subtle variations only emerge with a close reading of the text, and it would be easy to overlook them.
The published format of the texts also makes them appear similar in lettering and arrangement (e.g. left indentation).

@@

\begin{figure}[!hbp]
\centering
 \includegraphics[width=0.7\columnwidth]{PaperproposalforEAGLEfinal-img001.png}
\caption{Screenshot of IAph 4 webpage}
\label{fig:1}
\end{figure}

@@

\newpage
While traditional text-based modes of publication such as MAMA VIII have, at times, constrained the study of hands to an
analysis of lettering, digital corpora offer new opportunities to view, analyse and incorporate broader scope of visual
elements in the interpretation of an inscription. This is not to say that the traditional methods or charts of
letterforms development should be abandoned, merely that the methodology could be expanded, as it has been in this case
study, to include aspects of textual organization (spacing, use of decoration and punctuation, letter size). The
question arises: how do we achieve this through the website materials?

### 2016-02-10 source/22/Iscrizioni Latine Arcaiche - EAGLE 2016.tex
orfans and widows
percorsi tutti i file per eliminare orfane e vedove inserendo \newpage
o \vspace per spostare leggermente i paragrafi

The charts present several photographs of the inscriptions, taken during the photographic campaign carried out by the
project's team. Enlargements of some useful or problematic details of the inscription have been added: the richness in
images is related to the participation of the project in the Europeana network, a database of images of the European
cultural heritage. Although the images illustrate the inscription in an optimal way, fac-similes of each text will also
be provided, clearly related with the transcription provided by the editor.

We offer two kinds of transcription, the ‘Diplomatic transcription' and the ‘Interpretative transcription', which
provides the edition of the text. The need for a diplomatic transcription is due to the problematic nature of archaic
texts: in the case of the \emph{Duenos} vase, as it is well known, the second section of the text is almost always given by
scholars in diplomatic transcription because of the difficulties in segmenting the phrases into words (but with
attempts of interpretations of small portions).

>>>>

The charts present several photographs of the inscriptions, taken during the photographic campaign carried out by the
project's team. Enlargements of some useful or problematic details of the inscription have been added: the richness in
images is related to the participation of the project in the Europeana network, a database of images of the European
cultural heritage. Although the images illustrate the inscription in an optimal way, fac-similes of each text will also
be provided, clearly related with the transcription provided by the editor. 
\newpage
We offer two kinds of transcription, the ‘Diplomatic transcription' and the ‘Interpretative transcription', which
provides the edition of the text. The need for a diplomatic transcription is due to the problematic nature of archaic
texts: in the case of the \emph{Duenos} vase, as it is well known, the second section of the text is almost always given by
scholars in diplomatic transcription because of the difficulties in segmenting the phrases into words (but with
attempts of interpretations of small portions).

### 2016-02-10 source/28/EAGLE2016_Rocco_engrev.tex
further changes to verge, Graham and fixes to Vassallo
orfans and widows
percorsi tutti i file per eliminare orfane e vedove inserendo \newpage
o \vspace per spostare leggermente i paragrafi

%%%%%%%%
%TITLE:%
%%%%%%%%
\title{EDB 2.0 
How EAGLE improved the Epigraphic Database Bari}
\author[uniba]{Anita Rocco \corref{first}}
\address[uniba]{Dip. di Studi Umanistici, Università degli Studi di Bari ``Aldo Moro''}
\cortext[first]{Corresponding author. Email: anita.rocco@uniba.it}
\begin{document}
\maketitle
\begin{abstract}
This paper is dedicated to the evolution of the Epigraphic Database Bari (EDB) from the minimalistic design of its
origins to its current status. Although EDB, the database of inscriptions by Christians from Rome, dates back to the
late 1980s , involvement in the EAGLE – Europeana project has had a significantly positive impact on its development.
In fact, maintaining its peculiar character, dictated by its own history and, mostly, by the characteristics of its
documentary base, it has taken advantage of the solutions adopted to integrate different archives and purpose-built
best practices. 
\end{abstract}
\keywords{Epigraphic database, EDB, EAGLE Europeana, Christian inscriptions, Late antique inscriptions }

@@

Obviously this step has resulted in a series of changes and adjustments that led from the original basic structure of
the database to the present one.

@@

\newpage
\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{Fig4.png}
\caption{Bibliographic data fields.}
\label{fig:3b}
\end{figure}

@@

Is worth noting that this set of associated fields refers to the original position of the inscriptions and not to the
place where they have been found, unless the two data coincide, that is in the case of inscriptions \emph{in situ} or \emph{suo loco
adplicitae}, in accordance with the definitions of the ICVR.

To complete information on spatial data, all cemeterial contexts have been georeferenced, so that clicking on their name
opens a new window in \emph{Google Map} that shows the modern entrance to the cemetery, with the address and its coordinates.

@@

 \begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016Roccoengrev-img006.png} 
\caption{The original context maps.}
\label{fig:6}
\end{figure}

@@

\begin{figure}[hbp]
\centering
 \includegraphics[width=\columnwidth]{Fig7.png}
\caption{Conservation input fields.}
\label{fig:7}
\end{figure}

>>>>

%%%%%%%%
%TITLE:%
%%%%%%%%
\title{EDB 2.0.
How EAGLE improved the Epigraphic Database Bari}
\author[uniba]{Anita Rocco \corref{first}}
\address[uniba]{Dip. di Studi Umanistici, Università degli Studi di Bari ``Aldo Moro''}
\cortext[first]{Corresponding author. Email: anita.rocco@uniba.it}
\begin{document}
\maketitle
\begin{abstract}
This paper is dedicated to the evolution of the Epigraphic Database Bari (EDB) from the minimalistic design of its
origins to its current status. Although EDB, the database of inscriptions by Christians from Rome, dates back to the
late 1980s , involvement in the EAGLE – Europeana project has had a significantly positive impact on its development.
In fact, maintaining its peculiar character, dictated by its own history and, mostly, by the characteristics of its
documentary base, it has taken advantage of the solutions adopted to integrate different archives and purpose-built
best practices. 
\end{abstract}
\keywords{Epigraphic database, EDB, EAGLE Europeana, Christian inscriptions, Late antique inscriptions }

@@

\newpage
Obviously this step has resulted in a series of changes and adjustments that led from the original basic structure of
the database to the present one.

@@

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{Fig4.png}
\caption{Bibliographic data fields.}
\label{fig:3b}
\end{figure}

@@

Is worth noting that this set of associated fields refers to the original position of the inscriptions and not to the
place where they have been found, unless the two data coincide, that is in the case of inscriptions \emph{in situ} or \emph{suo loco
adplicitae}, in accordance with the definitions of the ICVR.
\newpage
To complete information on spatial data, all cemeterial contexts have been georeferenced, so that clicking on their name
opens a new window in \emph{Google Map} that shows the modern entrance to the cemetery, with the address and its coordinates.

@@

 \begin{figure}[!hbp]
\centering
 \includegraphics[width=0.9\columnwidth]{EAGLE2016Roccoengrev-img006.png} 
\caption{The original context maps.}
\label{fig:6}
\end{figure}

@@

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{Fig7.png}
\caption{Conservation input fields.}
\label{fig:7}
\end{figure}

### 2016-02-10 source/3/EAGLE_2016_submission_3.tex
further changes
Bronx not moved to pdfs folder, as it is being compiled by Francesco
fixes to Varga e Moucunill
further changes to verge, Graham and fixes to Vassallo
orfans and widows
percorsi tutti i file per eliminare orfane e vedove inserendo \newpage
o \vspace per spostare leggermente i paragrafi

\section{Motivation}

@@

@@

The construction of the database will follow a series of steps imposed by best practices: creating a repository of
sources; introducing, integrating, standardizing, coding and storing the information; enriching and disseminating the
information. The codifications hold an essential part, not only in the individual linkage procedures, but in the
analysing process as well. At the point when the database will comprise enough data, properly recorded and with all
codifications undertaken, the usage of statistical software in order to identify trends and run comparisons over large
scale geographical and administrative units might result in a better than ever understanding of Rome’s
social history.

\begin{figure}[!bp]
\centering
\includegraphics[scale=0.50]{1.png}
\caption{Schema of the database}
\label{fig:schemadatabase}
\end{figure}

@@

\noindent Within this theoretical frame, whose amplitude implies a gradual and long-term approach, we have chosen to start by
building the epigraphical division of the database. Thus, we began developing a project on the middle classes of the
Low Danube provinces and their epigraphic attestations. The database created in this context together with the platform
\url{www.romans1by1.com} represent the skeleton of the future population database of the Roman Empire.

For now, we are solely focusing on characters attested epigraphically – thus on inscriptions as sources. As constructing
a metadata suitable for all social and professional categories of the provincial world is very complex, our database is
created for accommodating all information epigraphically provided on members of the provincial middle classes.
Terminologically, we consider all those who manifest themselves epigraphically, without being members of the imperial
or provincial elite, as member of the middle socio-economical layers. Likewise, we have excluded active militaries (but
not their families), as their appurtenance to the army creates their social status.

Regarding the data entry, we are trying to remain faithful to the source and to record, during the first phase, only the
minimum of deduced information (gender, juridical status, ethnicity of the name). At the same time, we operate some
conventional onomastic transcriptions (AEL will be \emph{Aelius} from the start, etc.). All these basic normalization
procedures are being thoroughly explained and documented in the data entry manual, and the reason why we chose to apply
them is to speed up data standardization, run analyses and publish results on small samples, whose standardization
procedures would otherwise be overwhelmingly time-consuming in relation with the results.

@@

The \emph{Occupation} field requires some special attention, as it has the \emph{Occupation code} associated with it; as we are trying
to propose a codification of Roman occupations/professions, based on and adapting the HISCO classification
model\footnote{\url{http://historyofwork.iisg.nl/} }. While a raw classification and codification based on HISCO might only
be a slight challenge, finding a theoretical model might prove to be a more elaborated task. What we aim at is
constructing the ``metadata'' for the codification of Roman professions attested on stone and analyse how deep the
classification can realistically go. Much alike other normalization fields present in the sources component of the
database, \emph{Occupation code} was implemented because it helps dealing faster with small and medium size samples (up to
hundreds of people) in view of publishing preliminary results which are vital for dissemination and further financial
support of the project.

Two important variables, for statistics and working with the data, are \emph{Dedicated by} and \emph{Dedicated for}, which state the
relation of the recorded individual with the epigraph and with other persons mentioned by it. Both appear in the simple
form of check boxes.

Another particular check box, in need of supplementary explanations, is Later. The database was conceived for attested
civilians from the middle classes, but sometimes they are associated in inscriptions with militaries or representatives
of the elites. In this case, we have to register a minimum of data on the later as well, in order to build a wider
image of the characters that we focus on. When checking \emph{Later}, it opens \emph{Status}, which at its turn opens the following
options in the form of check boxes: \emph{Senator}, \emph{Knight}, \emph{Local magistrate}, \emph{Decurion}, \emph{Imperial priest}, \emph{Imperial slave},
\emph{Imperial freedman}, \emph{Military personnel}. If checked, each of these boxes opens a Details text box; additionally, \emph{Local
magistrate} and \emph{Decurion} open a \emph{City/Town} value list, while \emph{Military personnel}, opens \emph{Rank} and \emph{Unit} value lists. While
elite members and military personnel will at some point be added to the database, for the currently running project
purposes, their social and professional status, together with a minimum of relevant details and relation with the
recorded individuals represent enough information. 
\vspace{20pt}

%\begin{table}[!h]
\tablefirsthead{\hline \textbf{Field label} & \textbf{Field id} & \textbf{Data type} \\ \hline }
\bottomcaption{Metadata of the Personal data table}

@@

{\small
\addtolength{\tabcolsep}{-0.5mm}
%\begin{supertabular*}{\textwidth}{p{3.5cm}lp{3cm}}
\begin{xtabular}{|p{.30\textwidth}|p{.30\textwidth}|p{.30\textwidth}|}
%\begin{tabular*}{p{.30\textwidth}p{.29\textwidth}p{.31\textwidth}}
%\toprule
%\textbf{Field label} & \textbf{Field id} & \textbf{Data type}\\\midrule
Praenomen &
rperson\_praenomen &
Text\\ \hline
Ethnicity praenomen &
rperson\_eth\_praen &
Value list\\ \hline
Nomen &
rperson\_nomen &
Text\\ \hline
Ethnicity nomen &
rperson\_eth\_nom &
Value list\\ \hline
Cognomen/Personal name &
rperson\_cognomen &
Text\\ \hline
Ethnicity cognomen/personal name &
rperson\_eth\_cogn &
Value list\\ \hline
Father/Master name &
rperson\_father &
Text\\ \hline
Ethnicity father/master name &
rperson\_eth\_fth &
Value list\\ \hline
Agnomen &
rperson\_agnomen &
Text\\ \hline
Ethnicity agnomen &
rperson\_eth\_agn &
Value list\\ \hline
Observations agnomen &
rperson\_remark\_agn &
Text\\ \hline
Signum &
rperson\_signum &
Text\\ \hline
Ethnicity signum &
rperson\_eth\_sign &
Value list\\ \hline
Observations signum &
rperson\_remark\_sign &
Text\\ \hline
Natione &
rperson\_nation &
Text\\ \hline
Ethnicity &
rperson\_nat\_ethnicity &
Text\\ \hline
Gender &
rperson\_rgender\_id &
Value list\\ \hline
Juridical status &
rperson\_jstatus &
Checkbox. \ Opens checkboxes:

@@

This work began being developed during the period of a residential scholarship at the Hardt Fondation, Vandoeuvre. Its
expansion to the area of the whole Latin language Empire is being undertaken with the support of a postdoctoral
fellowship of the Fritz Thyssen Stiftung.

>>>>

\vspace{10pt}
\section{Motivation}

@@

\begin{figure}[!hbp]
\centering
\includegraphics[width=\columnwidth]{graph.png}
\caption{Schema of the database}
\label{fig:schemadatabase}
\end{figure}

@@

\newpage
The construction of the database will follow a series of steps imposed by best practices: creating a repository of
sources; introducing, integrating, standardizing, coding and storing the information; enriching and disseminating the
information. The codifications hold an essential part, not only in the individual linkage procedures, but in the
analysing process as well. At the point when the database will comprise enough data, properly recorded and with all
codifications undertaken, the usage of statistical software in order to identify trends and run comparisons over large
scale geographical and administrative units might result in a better than ever understanding of Rome’s
social history.

@@

\noindent Within this theoretical frame, whose amplitude implies a gradual and long-term approach, we have chosen to start by
building the epigraphical division of the database. Thus, we began developing a project on the middle classes of the
Low Danube provinces and their epigraphic attestations. The database created in this context together with the platform
\url{www.romans1by1.com} represent the skeleton of the future population database of the Roman Empire.
For now, we are solely focusing on characters attested epigraphically – thus on inscriptions as sources. As constructing
a metadata suitable for all social and professional categories of the provincial world is very complex, our database is
created for accommodating all information epigraphically provided on members of the provincial middle classes.
Terminologically, we consider all those who manifest themselves epigraphically, without being members of the imperial
or provincial elite, as member of the middle socio-economical layers. Likewise, we have excluded active militaries (but
not their families), as their appurtenance to the army creates their social status.
Regarding the data entry, we are trying to remain faithful to the source and to record, during the first phase, only the
minimum of deduced information (gender, juridical status, ethnicity of the name). At the same time, we operate some
conventional onomastic transcriptions (AEL will be \emph{Aelius} from the start, etc.). All these basic normalization
procedures are being thoroughly explained and documented in the data entry manual, and the reason why we chose to apply
them is to speed up data standardization, run analyses and publish results on small samples, whose standardization
procedures would otherwise be overwhelmingly time-consuming in relation with the results.

@@

The \emph{Occupation} field requires some special attention, as it has the \emph{Occupation code} associated with it; as we are trying
to propose a codification of Roman occupations/professions, based on and adapting the HISCO classification
model\footnote{\url{http://historyofwork.iisg.nl/} }. While a raw classification and codification based on HISCO might only
be a slight challenge, finding a theoretical model might prove to be a more elaborated task. What we aim at is
constructing the ``metadata'' for the codification of Roman professions attested on stone and analyse how deep the
classification can realistically go. Much alike other normalization fields present in the sources component of the
database, \emph{Occupation code} was implemented because it helps dealing faster with small and medium size samples (up to
hundreds of people) in view of publishing preliminary results which are vital for dissemination and further financial
support of the project. 
\newpage
Two important variables, for statistics and working with the data, are \emph{Dedicated by} and \emph{Dedicated for}, which state the
relation of the recorded individual with the epigraph and with other persons mentioned by it. Both appear in the simple
form of check boxes.

Another particular check box, in need of supplementary explanations, is Later. The database was conceived for attested
civilians from the middle classes, but sometimes they are associated in inscriptions with militaries or representatives
of the elites. In this case, we have to register a minimum of data on the later as well, in order to build a wider
image of the characters that we focus on. When checking \emph{Later}, it opens \emph{Status}, which at its turn opens the following
options in the form of check boxes: \emph{Senator}, \emph{Knight}, \emph{Local magistrate}, \emph{Decurion}, \emph{Imperial priest}, \emph{Imperial slave},
\emph{Imperial freedman}, \emph{Military personnel}. If checked, each of these boxes opens a Details text box; additionally, \emph{Local
magistrate} and \emph{Decurion} open a \emph{City/Town} value list, while \emph{Military personnel}, opens \emph{Rank} and \emph{Unit} value lists. While
elite members and military personnel will at some point be added to the database, for the currently running project
purposes, their social and professional status, together with a minimum of relevant details and relation with the
recorded individuals represent enough information. 
\newpage

%\begin{table}[!h]
\tablehead{\hline \textbf{Field label} & \textbf{Field id} & \textbf{Data type} \\ \hline}

{\small
\addtolength{\tabcolsep}{-0.5mm}
%\begin{supertabular*}{\textwidth}{p{3.5cm}lp{3cm}}
\begin{xtabular}{|p{.30\textwidth}|p{.33\textwidth}|p{.33\textwidth}|}

\tablefirsthead{\hline \textbf{Field label} & \textbf{Field id} & \textbf{Data type} \\ \hline }
\bottomcaption{Metadata of the Personal data table}

@@

%\begin{tabular*}{p{.30\textwidth}p{.29\textwidth}p{.31\textwidth}}
%\toprule
%\textbf{Field label} & \textbf{Field id} & \textbf{Data type}\\\midrule
Praenomen &
rperson\_praenomen &
Text\\ \hline
Ethnicity praenomen &
rperson\_eth\_praen &
Value list\\ \hline
Nomen &
rperson\_nomen &
Text\\ \hline
Ethnicity nomen &
rperson\_eth\_nom &
Value list\\ \hline
Cognomen/Personal name &
rperson\_cognomen &
Text\\ \hline
Ethnicity cognomen/personal name &
rperson\_eth\_cogn &
Value list\\ \hline
Father/Master name &
rperson\_father &
Text\\ \hline
Ethnicity father/master name &
rperson\_eth\_fth &
Value list\\ \hline
Agnomen &
rperson\_agnomen &
Text\\ \hline
Ethnicity agnomen &
rperson\_eth\_agn &
Value list\\ \hline
Observations agnomen &
rperson\_remark\_agn &
Text\\ \hline
Signum &
rperson\_signum &
Text\\ \hline
Ethnicity signum &
rperson\_eth\_sign &
Value list\\ \hline
Observations signum &
rperson\_remark\_sign &
Text\\ \hline
Natione &
rperson\_nation &
Text\\ \hline
Ethnicity &
rperson\_nat\_ethnicity &
Text\\ \hline
Gender &
rperson\_rgender\_id &
Value list\\ \hline
Juridical status &
rperson\_jstatus &
Checkbox. \ Opens checkboxes:

@@
